\bf Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-01-19 paper.qmd
Add remove low density hex, long edges MSE, AIC plots with different evaluations

df_bin_n <- df_bin |>
  mutate(ID = 1:NROW(df_bin))

tr_from_to_df_n <- left_join(tr_from_to_df, df_bin_n, by = c("from" = "ID")) |>
  rename("from_hb_id" = "hb_id")

tr_from_to_df_n <- left_join(tr_from_to_df_n, df_bin_n, by = c("to" = "ID")) |>
  rename("to_hb_id" = "hb_id")

hb_id_selected <- tr_from_to_df_n |>
  dplyr::filter((from == 2) & (to == 3)) |>
  head(1)

@@

tr_from_to_df_coord_with_group <- merge(tr_from_to_df, distance_edges, by = c("from", "to"))

@@

# bin_coord_1 <- tr_df |>
#   dplyr::filter((x == hb_id_selected$x_from)) |>
#   dplyr::filter((y == hb_id_selected$y_from))

bin_coord_1 <- tr_df |>
  head(1)

# bin_coord_2 <- tr_df |>
#   dplyr::filter((x == hb_id_selected$x_to)) |>
#   dplyr::filter((y == hb_id_selected$y_to))

bin_coord_2 <- tr_df |>
  dplyr::filter(row_number() == 2)

@@

a1 <- ggplot(bin_coord, aes(x = x, y = y)) +
  geom_segment(
    aes(x = x_from, y = y_from, xend = x_to, yend = y_to),
    data = tr_from_to_df_coord_with_group, colour = "#f6e8c3"
  ) +
  geom_point() +
  coord_equal() +
  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), alpha = 0.5, size = 1) +
  scale_color_manual(values=c("#e6550d", "#000000")) +
  geom_point(size = 3, colour = "#33a02c") + 
  #ggtitle("(b)") + 
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
  #coord_equal() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  )

@@

```{r}
#| warning: false
#| echo: false
#| label: fig-wkhighD
#| fig-cap: How the 2D model lift into high dimensions? (a) visualize the points and the hexagonal bin centroids related 2nd and 15th hexagons, (b) visualization of the edge connected the 2nd and 15th hexagons (colored in red) in the triangular mesh. A video of tour animation is available at <https://youtu.be/yN-RNPV_cHQ>.

@@

- Compute high-D distances (using averaged high-D data)
- Fit a linear model using response as distance_high_D and predictor as distance_2D
- Filter the model residuals only for short edges
- compute error for each edge connection as (residual for that point)/(total residual for all edges * number of short edges)

error_plot_lg_1 <- ggplot(error_df_lg_1, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_colour_manual(values = c("#e41a1c")) +
  ylab("error") +
  xlab("Benchmark value")

error_plot_lg_2 <- ggplot(error_df_lg_2, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_colour_manual(values = c("#377eb8")) +
  ylab("error") +
  xlab("Benchmark value")

error_plot_lg_3 <- ggplot(error_df_lg_3, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_colour_manual(values = c("#4daf4a")) +
  ylab("error") +
  xlab("Benchmark value")

error_plot_lg_4 <- ggplot(error_df_lg_4, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_colour_manual(values = c("#984ea3")) +
  ylab("error") +
  xlab("Benchmark value")

error_plot_lg_5 <- ggplot(error_df_lg_5, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_colour_manual(values = c("#ff7f00")) +
  ylab("error") +
  xlab("Benchmark value")
```

error_plot_lg_1 + error_plot_lg_2 + error_plot_lg_3 + error_plot_lg_4 + error_plot_lg_5 +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect', nrow = 2) &
  theme(legend.position='bottom', plot.tag = element_text(size = 8)) 
```

@@

>>>>

df_bin_centroids_n <- df_bin_centroids |>
  select(hexID) |>
  mutate(ID = 1:NROW(df_bin_centroids))

tr_from_to_df_n <- left_join(tr_from_to_df, df_bin_centroids_n, by = c("from" = "ID")) |>
  rename("from_hb_id" = "hexID")

tr_from_to_df_n <- left_join(tr_from_to_df_n, df_bin_centroids_n, by = c("to" = "ID")) |>
  rename("to_hb_id" = "hexID")

hb_id_selected <- tr_from_to_df_n |>
  dplyr::filter((from == 2) & (to == 3))

@@

tr_from_to_df_coord_with_group <- inner_join(tr_from_to_df, distance_edges, by = c("from" = "from", "to" = "to"))

@@

bin_coord_1 <- tr_df |>
  dplyr::filter((x == hb_id_selected$x_from)) |>
  dplyr::filter((y == hb_id_selected$y_from))

# bin_coord_1 <- tr_df |>
#   head(1)

bin_coord_2 <- tr_df |>
  dplyr::filter((x == hb_id_selected$x_to)) |>
  dplyr::filter((y == hb_id_selected$y_to))

# bin_coord_2 <- tr_df |>
#   dplyr::filter(row_number() == 2)

@@

a1 <- ggplot(bin_coord, aes(x = x, y = y)) +
  geom_segment(
    aes(x = x_from, y = y_from, xend = x_to, yend = y_to, colour = type),
    data = tr_from_to_df_coord_with_group) +
  geom_point() +
  coord_equal() +
  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), alpha = 0.5, size = 1) +
  scale_color_manual(values=c("#e6550d", "#000000", "#e31a1c", "#f6e8c3")) +
  geom_point(size = 3, colour = "#33a02c") + 
  #ggtitle("(b)") + 
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
  #coord_equal() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  )

@@

```{r}
#| warning: false
#| echo: false
#| label: fig-wkhighD
#| fig-cap: How the 2D model lift into high dimensions? (a) visualize the points and the hexagonal bin centroids related 2nd and 15th hexagons, (b) visualization of the edge connected the 2nd and 15th hexagons (colored in red) in the triangular mesh. A video of tour animation is available at <https://youtu.be/ffWFVK1UTIc>.

@@

@@

## To draw with AIC
aic_plot_lg <- ggplot(error_df_lg, aes(x = benchmark_rm_lg,
                                                                                 y = total_error,
                                                                                 color = method
)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
  ylab("AIC") +
  xlab("Benchmark value")
## Effective number of bins along x-axis

mse_plot_lg <- ggplot(error_df_lg, aes(x = benchmark_rm_lg,
                                       y = total_mse,
                                       color = method
)) +
  geom_point() +
  geom_line() +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
  ylab("MSE") +
  xlab("Benchmark value")

aic_plot_lg + mse_plot_lg +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect', ncol = 2) &
  theme(legend.position='bottom', plot.tag = element_text(size = 8)) 
```

### 2024-01-19 paper.tex
Add remove low density hex, long edges MSE, AIC plots with different evaluations

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, enhanced, sharp corners, breakable]}{\end{tcolorbox}}\fi

@@

\caption{\label{fig-wkhighD}How the 2D model lift into high dimensions?
(a) visualize the points and the hexagonal bin centroids related 2nd and
15th hexagons, (b) visualization of the edge connected the 2nd and 15th
hexagons (colored in red) in the triangular mesh. A video of tour
animation is available at \url{https://youtu.be/yN-RNPV_cHQ}.}

@@

\begin{itemize}
\tightlist
\item
  Compute high-D distances (using averaged high-D data)
\item
  Fit a linear model using response as distance\_high\_D and predictor
  as distance\_2D
\item
  Filter the model residuals only for short edges
\item
  compute error for each edge connection as (residual for that
  point)/(total residual for all edges * number of short edges)
\end{itemize}

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, breakable, enhanced, sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi

@@

\caption{\label{fig-wkhighD}How the 2D model lift into high dimensions?
(a) visualize the points and the hexagonal bin centroids related 2nd and
15th hexagons, (b) visualization of the edge connected the 2nd and 15th
hexagons (colored in red) in the triangular mesh. A video of tour
animation is available at \url{https://youtu.be/ffWFVK1UTIc}.}

@@
